Visualizing predictions

### 2021-02-13 effects_writeup.Rnw
Autocommit
Added contrast

#library(shellpipes)
#sourceFiles()

@@

>>>>

load("predict_plots.rda")

## How to pass arguments??
#library(shellpipes)
#sourceFiles()

@@

\section{Unbalanced data}

Consider a simple no-interaction terms simulation:
\begin{itemize}
\item $y = \beta_0 + \beta_1x_1 + \beta_2x_2 + \epsilon$ s.t $\epsilon \sim N(0, 1)$ and $\{\beta_0 = -5, \beta_1 = 0.1, \beta_2 = \{-0.6, 0.5\}\}$
\begin{itemize}
\item $x_1 \sim unif(1, 9)$
\item $x_2$ -- three level categorical variable s.t $A$, $B$ and $C$ are $10\%$, $30\%$ and $60\%$, respectively.
\end{itemize}
\end{itemize}

<<simple_sim, echo=FALSE>>=
set.seed(101)
N <- 500
x1_min <- 1
x1_max <- 9
b0 <- -0.5
b_x1 <- 0.1
b_x2b <- -0.6
b_x2c <- 0.5

cat_props <- c(0.1, 0.3, 0.6)
df <- data.frame(x1u = runif(n=N, min=x1_min, max=x1_max)
                  , x2u = sample(c("A", "B", "C"), N, prob = cat_props, replace = TRUE)
)
X <- model.matrix(~x1u + x2u, df)
betas <- c(b0, b_x1, b_x2b, b_x2c)
df$y <- rnorm(nrow(df), mean= X %*% betas, sd=1)
df <- transform(df
	, x2c = x2u
	, x1s = drop(scale(x1u, center=FALSE))
	, x1sd = drop(scale(x1u))
)
prop.table(table(df$x2u))
head(df)
@

We fit two models -- the model without contrast and the one with sum-to-contrast (\code{`contr.sum'}):

<<contrast_modes>>=
## No contrast model
lm_no_contr <- lm(y ~ x1u + x2u, data = df)
print(lm_no_contr)

## With contrast
contrasts(df$x2c) <- "contr.sum"
lm_contr <- lm(y ~ x1u + x2c, data = df)
print(lm_contr)
@

First, we include the intercept variance in the variable effects:

<<simple_model, echo=FALSE>>=
contr_models <- list(x2u = lm_no_contr, x2c = lm_contr)
@

<<contr_model_predictions, echo=FALSE, results=hide>>=
## varpred
contr_vpred_all <- condsummary(mod_list=contr_models, fun=function(x, f){
	dd <- varpred(x, f)
}, simplify = TRUE, combine = "rbind", fmethod = "varpred")

## emmeans
contr_empred_all <- condsummary(mod_list=contr_models, fun=function(x, f){
	spec <- as.formula(paste0("~", f))
	dd <- emmeans(x, spec=spec, cov.keep=f)
}, simplify = TRUE, combine = "rbind", fmethod = "emmeans")

## effects
contr_efpred_all <- condsummary(mod_list=contr_models, fun=function(x, f){
	dd <- Effect(f, x)
}, simplify = TRUE, combine = "rbind", fmethod = "effects")

## Combine all the estimates
contr_pred_all <- do.call("rbind"
	, list(contr_vpred_all, contr_empred_all, contr_efpred_all)
)
@

<<echo=FALSE>>=
class(contr_pred_all) <- c("jdeffects", "data.frame")
contr_pred_all_plot_out <- (plot(contr_pred_all)
	+ facet_wrap(~method)
	+ theme(legend.position="bottom")
)
@

\begin{figure}[H]
\begin{center}
<<contr_model_predictions_plot, fig=TRUE, echo=FALSE>>=
contr_pred_all_plot_out
@
\end{center}
\caption{Variable effects with intercept and other non-focal predictor uncertainties included. The legend compares the model with non contrast (\code{x2u}) to the one with sum to zero contrast (\code{x2c}).}
\label{fig:contr_pred_plot}
\end{figure}

As explained in the previous section, we again eliminate the uncertainties due to the intercept term and the non-focal variables.

<<contr_model_zeroed_vcov, echo=FALSE, results=hide>>=
## varpred
contr_vcov_vpred_all <- condsummary(mod_list=contr_models, fun=function(x, f){
	dd <- varpred(x, f, vcov. = zero_vcov(x, f))
}, simplify = TRUE, combine = "rbind", fmethod = "varpred")

## emmeans
contr_vcov_empred_all <- condsummary(mod_list=contr_models, fun=function(x, f){
	spec <- as.formula(paste0("~", f))
	dd <- emmeans(x, spec=spec, cov.keep=f, vcov. = zero_vcov(x, f))
}, simplify = TRUE, combine = "rbind", fmethod = "emmeans")

## effects
contr_vcov_efpred_all <- condsummary(mod_list=contr_models, fun=function(x, f){
	dd <- Effect(f, x, vcov. = function(x, complete=FALSE)zero_vcov(x, f, complete))
}, simplify = TRUE, combine = "rbind", fmethod = "effects")

## Combine all the estimates
contr_vcov_pred_all <- do.call("rbind"
	, list(contr_vcov_vpred_all, contr_vcov_empred_all, contr_vcov_efpred_all)
)
@

<<echo=FALSE>>=
class(contr_vcov_pred_all) <- c("jdeffects", "data.frame")
contr_pred_zeroed_vcov_plot <- (plot(contr_vcov_pred_all)
	+ facet_wrap(~method)
	+ theme(legend.position="bottom")
)
@

\begin{figure}[H]
\begin{center}
<<contr_zeroed_vcov_plot, fig=TRUE, echo=FALSE>>=
contr_pred_zeroed_vcov_plot
@
\end{center}
\caption{Variable effects with intercept and other non-focal predictor uncertainties included. The legend compares the model with non contrast (\code{x2u}) to the one with sum to zero contrast (\code{x2c}).}
\label{fig:contr_zeroed_vcov}
\end{figure}

<<contr_model_zeroed_mmat, echo=FALSE, results=hide>>=
## varpred
contr_mmat_vpred_all <- condsummary(mod_list=contr_models, fun=function(x, f){
	dd <- varpred(x, f, isolate = TRUE, vcov. = zero_vcov(x, f))
}, simplify = TRUE, combine = "rbind", fmethod = "varpred")

## emmeans
contr_mmat_empred_all <- condsummary(mod_list=contr_models, fun=function(x, f){
	spec <- as.formula(paste0("~", f))
	dd <- emmeans(x, spec=spec, cov.keep=f, vcov. = zero_vcov(x, f))
}, simplify = TRUE, combine = "rbind", fmethod = "emmeans")

## effects
contr_mmat_efpred_all <- condsummary(mod_list=contr_models, fun=function(x, f){
	dd <- Effect(f, x, vcov. = function(x, complete=FALSE)zero_vcov(x, f, complete))
}, simplify = TRUE, combine = "rbind", fmethod = "effects")

## Combine all the estimates
contr_mmat_pred_all <- do.call("rbind"
	, list(contr_mmat_vpred_all, contr_mmat_empred_all, contr_mmat_efpred_all)
)
@

<<echo=FALSE>>=
class(contr_mmat_pred_all) <- c("jdeffects", "data.frame")
contr_pred_zeroed_mmat_plot <- (plot(contr_mmat_pred_all)
	+ facet_wrap(~method)
	+ theme(legend.position="bottom")
)
@

\begin{figure}[H]
\begin{center}
<<contr_zeroed_mmat_plot, fig=TRUE, echo=FALSE>>=
contr_pred_zeroed_mmat_plot
@
\end{center}
\caption{Variable effects with intercept and other non-focal predictor uncertainties included. The legend compares the model with non contrast (\code{x2u}) to the one with sum to zero contrast (\code{x2c}).}
\label{fig:contr_zeroed_mmat}
\end{figure}
